### 2018-01-30 Chapters/6-TrackingOptimization.tex

\chapter{Tracking Optimization}
Your first chapter is probably an introduction. But who knows.