### 2018-02-04 Chapters/6-TrackingOptimization.tex

\chapter{Tracking Optimization}
Understanding how kaons and pions are tracked inside the TPC and optimizing the reconstruction algorithms to maximize the correct identification of the interaction point is a fundamental step of the analysis.

\section{MC sample and WC2TPC match}
The optimization is performed on a MC sample of \textcolor{red}{HOW MANY?} kaons and \textcolor{red}{HOW MANY?} pions produced with the DDMC technique. 
DDMC particles are shot from the WC4 location into the TPC following the beam profile.
We mimic the matching between the WC and the TPC track on Monte Carlo by constructing a fake WC track using truth information at the true initial point. We then apply the same WC to TPC matching algorithm as in data.

Plots I want in this section:
\begin{enumerate}
\item WC2TPC MC DeltaX, DeltaY and $\alpha$
\item Delta L, reco - true
\item Delta L, reco - true Elastic, Delta L, reco - true Inelastic, other
\item Length Quality cut
\item Efficiency as a function of true KE and Angle
\end{enumerate}