### 2018-05-14 Chapter-8/8-CrossSectionAnalysis.tex
s

\chapter{Negative Pion Cross Section Measurement}

\section{Estimate of $E_{loss}$ before the TPC}
\section{Interacting and Incident Distributions}

\section{Total Hadronic Negative Pion-Argon Differential Cross Section}