### 2018-05-25 thesis_main.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% yale_thesis.tex
% Elena Gramellini
% Started: 2017/12/05
%
% A bare, sample template for a Yale PhD thesis using yalephd.cls
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[letterpaper,12pt]{yalephd}
% remove draft option for final printing.
% font size must be between 10pt-12pt.
\usepackage{caption,setspace}
\usepackage{amsmath} 
\usepackage{rotating}

\usepackage{geometry} % you need this for yalephd.cls to work.
\usepackage{graphicx}
\usepackage[graphicx]{realboxes}
 % you probably want the rest of these.
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{hyperref}
\usepackage[export]{adjustbox}

\usepackage{mathtools}
\DeclarePairedDelimiter\bra{\langle}{\rvert}
\DeclarePairedDelimiter\ket{\lvert}{\rangle}
\DeclarePairedDelimiterX\braket[2]{\langle}{\rangle}{#1 \delimsize\vert #2}
\usepackage{textgreek}

\usepackage{multirow}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{appendix}
\usepackage{comment}
\usepackage{cite}
\usepackage{color}
\bibliographystyle{plain}
\usepackage{slashed}

%\setcounter{chapter}{-1}

\newenvironment{dedication}
  {%\clearpage           % we want a new page          %% I commented this
   \thispagestyle{empty}% no header and footer
   \vspace*{\stretch{1}}% some space at the top
   \itshape             % the text is in italics
   \raggedleft          % flush to the right margin
  }
  {\par % end the paragraph
   \vspace{\stretch{3}} % space at bottom is three times that at the top
   \clearpage           % finish off the page
  }

\usepackage{lineno,xcolor}

\linenumbers

\begin{document}

% Need to define title before the abstract.
\title{Measurement of ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross sections in the LArIAT experiment}
\author{Elena Gramellini}
\advisor{Bonnie T. Fleming}
\date{Date you'll receive your degree} % usually not \today.

% All the stuff at the front of your thesis.
\frontmatter

\begin{abstract}
Abstract goes here. Limit 750 words.
\end{abstract}

\maketitle
\makecopyright{2018}

 \begin{dedication}
A mia mamma e mio babbo,\\
grazie per le radici e grazie per le ali.
    \par   %% or a blank line
    \vspace{2\baselineskip}
To my mom and dad,\\
thank you for the roots and thank you for the wings.
    \vspace{\baselineskip}
  \end{dedication}

\tableofcontents
%\listoffigures % remove this if you have no figures.
%\listoftables % remove this if you have no tables.

\chapter{Acknowledgements} % this needs to be before \mainmatter.

{\raggedleft ``\emph{Dunque io ringrazio tutti quanti.} \par}
{\raggedleft \emph{Specie la mia mamma che mi ha fatto cos\'i funky.}"\par}
{\raggedleft -- Articolo 31, Tanqi Funky, 1996 -- \par}
\vspace{0.5cm}

{\raggedleft ``\emph{At last, I thank everyone.} \par}
{\raggedleft \emph{Especially my mom who made me so funky.}"\par}
{\raggedleft -- Articolo 31, Tanqi Funky, 1996 -- \par}
\vspace{0.5cm}

A lot of people are awesome, especially you, since you probably agreed to read this when it was a draft.

I jumped the pond a little less than 5 years ago because I had a taste of what being a physicists in the US might mean and wanted more. 
I came to this vast (and at times inhospitable) land thinking I would learn a lot of physics (which I did) and trying to keep doing research for as long as I could (which I'm fortunate enough to say I'll still do for a while). Little did I know that getting my PhD would be first and foremost a journey to self discovery, the beginning of a path to asymptotically tend to the person and the scientist I want to be. A journey impossible without the help of many people, which I will now proceed to thank, hoping that the acknowledgements won't be longer than the thesis itself.

Mom, you are and will always be the hero of my story; I know this has been challenging for you, thanks for being there for me at every hour of the day and the night, for being always honest with me and for somehow handling this whole US-Italy thing gracefully. I can't believe you still haven't met Bonnie. Ah, and don't worry, I'll always translate things for you.

Bonnie Fleming, my advisor, I don't even know where to start in thanking you. Thanks for listening to me before I proved I was worth be listened to; thanks for always believing I could make it, even at the bottom of my doubts; thanks for the laughters, thanks for leading by example, thanks for  letting me part of the most wonderful research group ;

%Flavio Cavanna 
%Ornella Palamara, I once told Andrzej I think of you as my spirit animal: fierce, independent and direct; he replied I still need to eat a lot of paninis to get there and I can't agree more.

%A big thank you to my academic sisters, from the youngest to the oldest (again academically speaking): Supraja Balasubramanian, you worked the magic of enriching my past, thank you from the bottom of my heart;  Brooke Morrison, the depth of your knowledge, your drive and love for our field has been and will always be a constant inspiration;  Ariana  Hackenburg, thanks for keeping me sane for the last four years, if you run for president -- as you should -- remember to delete our Skype chats, love and hugs to you dudette; and last but not least Xiao Luo, Fermilab has given me lots of extraordinary gifts, but no one  bigger than your friendship.

%Jen Raaf
%Jonathan Asaadi
%Andrzej M. Szelc
%Roberto Acciarri
%LArIAT & MicroBooNE
%To all the members of the NPG, thanks for all the laughters and all the wise insights. You're the smartest bunch of idiots I've ever met, I love y'all.
%Shany %Louise

%To the extraordinary people who stayed close, even with an ocean in between: la Tasso, Marci, la Bea, la Kittirn, Alvin, la Viola e Brogna, Sighi and all the people from the ``La Chiappia Nuova" group, thanks for making me feel like I never left.
%Finally, Anto, cause people might think I'm nice and easy, but you know I'm not. And you're still here.

% Starts proper arabic numbering of pages and chapters.
\mainmatter
\input{Chapter-0/0-Introduction}
\input{Chapter-1/1-Theory}
\input{Chapter-2/2-LArTPCs}
\input{Chapter-3/3-TheLArIATExperiment}
\input{Chapter-4/4-Methodology}
\input{Chapter-5/5-Preparatory}
\input{Chapter-6/6-PionCrossSection}
\input{Chapter-7/7-KaonCrossSection}
\input{Chapter-8/8-Conclusions}

%
%\input{Chapter-9/9-Uncertainty}

% Only call appendix once, here.
\appendix
%%\input{Chapters/A-KaonAna}
\input{AppendixB-TrackingStudies/B-Tracking}
\input{AppendixC-EnergyCalibration/C-EnergyCalibration}
%\input{AppendixB-EField/B-LArIATEField}
%%\input{Chapters/C-MicroBooNECRT}

% Old chapters configuration %
%\input{Chapters/4-DataCollection}
%\input{Chapters/4-MonteCarlo}
%\input{Chapters/5-EnergyCalibration}
%\input{Chapter-8/8-TrackingOptimization}

% Add additional \chapter{}s as necessary.

% use \cite{} to cite a reference in your bibliography file.
% use \ref{} to reference a \label{} from an equation, figure, or table.

% for sets of equations use align:
%\begin{align}
%\end{align}

% for figures:
%\begin{figure}[ht]
%\centering
%\includegraphics[width=.45\textwidth]{name_of_figure.eps}
%\caption{A caption! \label{a_figure}}
%\end{figure}

% for tables:
%\begin{table}
%\begin{tabular}{c|c|c}
% 1 & 2 & 3 \\
%\hline
%\end{tabular}
%\caption{Another caption! \label{a_table}}
%\end{table}

% Any chapters such as End Notes go after this.
\backmatter

\bibliography{bib.bib}
%\bibliographystyle{plain}
% for your own sake, use a bibtex file, so all of the numbering of references will be done
% automatically.

\end{document}