### 2018-05-30 thesis_main.tex
s

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% yale_thesis.tex
% Elena Gramellini
% Started: 2017/12/05
%
% A bare, sample template for a Yale PhD thesis using yalephd.cls
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[letterpaper,12pt]{yalephd}
% remove draft option for final printing.
% font size must be between 10pt-12pt.
\usepackage{caption,setspace}
\usepackage{amsmath} 
\usepackage{rotating}
%\usepackage{newtxtext,newtxmath,amsmath}
\usepackage{ textcomp }

\usepackage{geometry} % you need this for yalephd.cls to work.
\usepackage{graphicx}
\usepackage[graphicx]{realboxes}
 % you probably want the rest of these.
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{hyperref}
\usepackage[export]{adjustbox}

\usepackage{mathtools}
\DeclarePairedDelimiter\bra{\langle}{\rvert}
\DeclarePairedDelimiter\ket{\lvert}{\rangle}
\DeclarePairedDelimiterX\braket[2]{\langle}{\rangle}{#1 \delimsize\vert #2}
\usepackage{textgreek}

\usepackage{multirow}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{appendix}
\usepackage{comment}
\usepackage{cite}
\usepackage{color}
\bibliographystyle{plain}
\usepackage{slashed}

%\setcounter{chapter}{-1}

\newenvironment{dedication}
  {%\clearpage           % we want a new page          %% I commented this
   \thispagestyle{empty}% no header and footer
   \vspace*{\stretch{1}}% some space at the top
   \itshape             % the text is in italics
   \raggedleft          % flush to the right margin
  }
  {\par % end the paragraph
   \vspace{\stretch{3}} % space at bottom is three times that at the top
   \clearpage           % finish off the page
  }

\usepackage{lineno,xcolor}

\linenumbers

\begin{document}

% Need to define title before the abstract.
\title{Measurement of ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross sections in the LArIAT experiment}
\author{Elena Gramellini}
\advisor{Bonnie T. Fleming}
\date{Date you'll receive your degree} % usually not \today.

% All the stuff at the front of your thesis.
\frontmatter

\begin{abstract}
%Abstract goes here. Limit 750 words.

The Liquid Argon Time Projection Chamber (LArTPC) represents one of the
most advanced experimental technologies for physics at the Intensity
Frontier due to its full 3D-imaging, excellent particle identification
and precise calorimetric energy reconstruction. By deploying a LArTPC in
a dedicated calibration test beam line at Fermilab, LArIAT (Liquid
Argon In A Testbeam) aims to experimentally calibrate this technology
in a controlled environment and to provide physics results key to the
neutrino oscillation physics and proton decay searches of the Short
Baseline Neutrino and Long Baseline Neutrino programs.

LArIAT's physics program entails a vast set of topics with a
particular focus on the study of nuclear effects such as pion and
kaon characteristic interaction modes.

This thesis presents two world's first measurements: the measurement of ($\pi^-$-Ar) total hadronic cross section in the 100-1050 MeV kinetic energy range and the measurement of the ($K^+$-Ar) total hadronic cross section in the 100-650 MeV kinetic energy range. The analyses devised for these measurements use both the core elements of LArIAT: beamline and TPC. The first step in each analysis is the development of an event selection based on  beamline and TPC information geared towards the identification of the hadron of interest. We then proceed to match the beamline candidate to a suitable TPC track. Finally,  we apply the ``thin slice method" technique and measure the cross section, correcting for background and detector effects. The thin slice technique is a new method to measure hadron-argon cross sections possible only thanks to the combination of the tracking and calorimetry capability of the LArTPC technology. 
Albeit never on argon, the hadronic cross section of pions has been measured before on several different elements in thin target experiments, leading to solid predictions in the argon case. Through the use of a different technique, our measurement of the  ($\pi^-$-Ar) total hadronic cross section is in general agreement with the prediction by the Geant4 Bertini Cascade model which are based on data from thin target experiments. On the contrary, cross section measurements for kaons are extremely rare, thus more difficult to model. Not surprisingly, our measurement of the  ($K^+$-Ar) total hadronic cross section is mostly in disagreement with the Geant4 prediction.

This thesis also reports two  ancillary  detector physics measurements necessary for the cross section analyses: the measurements of the LArIAT electric field and calorimetry constants. We developed a technique to measure the LArIAT electric field using cathode-anode piercing tracks with cosmic data. We applied a new technique for the measurement of the calorimetry calibration constants based on the particles' momentum measurement.

The ($\pi^-$-Ar) and the ($K^+$-Ar) total hadronic cross measurements are the first physics results of the LArIAT experiment and will be the basis for the future LArIAT measurements of pion and kaon cross sections in the exclusive channels.

\end{abstract}

\maketitle
\makecopyright{2018}

 \begin{dedication}
A mia mamma e mio babbo,\\
grazie per le radici e grazie per le ali.
    \par   %% or a blank line
    \vspace{2\baselineskip}
To my mom and dad,\\
thank you for the roots and thank you for the wings.
    \vspace{\baselineskip}
  \end{dedication}

\tableofcontents
%\listoffigures % remove this if you have no figures.
%\listoftables % remove this if you have no tables.

\chapter{Acknowledgements} % this needs to be before \mainmatter.

{\raggedleft ``\emph{Dunque io ringrazio tutti quanti.} \par}
{\raggedleft \emph{Specie la mia mamma che mi ha fatto cos\'i funky.}"\par}
{\raggedleft -- Articolo 31, Tanqi Funky, 1996 -- \par}
\vspace{0.5cm}

{\raggedleft ``\emph{At last, I thank everyone.} \par}
{\raggedleft \emph{Especially my mom who made me so funky.}"\par}
{\raggedleft -- Articolo 31, Tanqi Funky, 1996 -- \par}
\vspace{0.5cm}

%A lot of people are awesome, especially you, since you probably agreed to read this when it was a draft.

%I jumped the pond a little less than 5 years ago because I had a taste of what being a physicists in the US might be like and wanted more. 
%I came to this vast and at times inhospitable land thinking I would learn a lot of physics (which I did) and hoping I'd keep doing research for as long as I could (which I'm fortunate enough to say I'll still do for a while). Little did I know that getting my PhD would be first and foremost a journey to self discovery, the beginning of a path to asymptotically tend to the person and the scientist I want to be. A journey impossible without the help of many people, which I will now proceed to thank, hoping that the acknowledgements won't be longer than the thesis itself.

Mom, you are and will always be the hero of my story; I know this has been challenging for you, thanks for being there for me at every hour of the day and the night, for being always honest with me and for somehow handling this whole US-Italy thing gracefully. I can't believe you still haven't met Bonnie. Ah, and don't worry, I'll always translate things for you.

Bonnie Fleming, my advisor, I don't even know where to start in thanking you. Thanks for listening to me before I proved I was worth be listened to; thanks for always believing I could make it, even at the bottom of my doubts; thanks for all the laughters, for all the freedom, for all the great advices; thanks for leading by example; thanks for being always supportive of all my crazy side projects; thanks for choosing each and every person in the most wonderful research group ever and thanks for letting me be part of it. Thanks for showing me that even Bonnie Freaking Fleming is human, has her downs and still she strives.

Flavio Cavanna, my co-advisor, working with you has been one of the greatest pleasures of my academic life; the extent of your knowledge never ceases to amaze me and the freshness of your enthusiasm is always contagious. Thanks for all the time spent in your office banging our heads on these cross section analyses and thanks for convincing me to join LArIAT in the first place.

Ornella Palamara, after a particularly heated exchange at a wine and cheese, I once told Andrzej I think of you as my spirit animal: fierce, knowledgeable and direct; he replied I still need to walk many miles to get there and I can't agree more.

A big thank you to my academic sisters, from the youngest to the oldest (again academically speaking): Supraja Balasubramanian, you worked the magic of enriching my past, thank you from the bottom of my heart;  Brooke Morrison, the depth of your knowledge, your drive and love for our field has been and will always be a constant inspiration;  Ariana  Hackenburg, thanks for keeping me sane for the last four years, if you run for president -- as you should -- remember to delete our Skype chats, love and hugs to you dudette; and last but not least Xiao Luo, Fermilab has given me lots of extraordinary gifts, but no one  bigger than your friendship.

Jen Raaf and Jonathan Asaadi, these analyses would have not seen the light of day without you. Thanks for your many inputs and for letting your door open to talk about physics and/or my latest paranoia. Thanks for the patience and understanding. And thanks for leading that small experiment with a big heart called LArIAT and all its crazy ponies. Speaking of which, thanks to all my collaborators, especially Irene Nutini, who started the path for the pion analysis, Greg Pulliam, who suffered with me at the end of said path, Dan the Man, Jason, Johnny, Will Fo and Mr Bojangles. A special token of gratitude goes to Roberto Acciarri and Andrzej M. Szelc, you are so much more to me than collaborators and previous car owners, you know you'll always have a home in Chicago.

To all the members of the NPG, thanks for all the laughters and all the wise insights. You're the smartest bunch of idiots I've ever met, I love y'all.

Thanks to the MicroBooNE collaboration Sam kazu and kirby

Shany Danieli and Louise Laage, you're the piece of heart I left in New Haven. Shany, I don't know what made you think I was the cool kid in the group, but I'm so glad you stuck with me. I'm not sure I would have survived my first semester at Yale without you: you shine in your smartness and lift other people with you -- it's a rare quality in our field. Louise, my dear, thanks for always finding the time to listen, for your attuned spirit and for the magnolias.

Thanks to all the people which helped me make Chicago and Fermilab my new home, in particular David Caratelli,  Mateus Carneiro, Pilar Coloma, Ciaran Hughes, Gordan Krnjaic, Seyda Ipek, Kiel Howe and Rob Fine.

To the extraordinary people who stayed close, even with an ocean in between: la Tasso, Marci, la Bea, la Kittirn, Alvin, la Viola e Brogna, Sighi, Eleonora and all the people from the ``La Chiappia Nuova" group, thanks for making me feel like I never left.

Finally, thanks to Anto, my husband, my topo and my adventure companion, cause people might think I'm nice and easy, but you know I'm not... And still, you followed me across half the world and you're still here, so that I could never feel alone.

% Starts proper arabic numbering of pages and chapters.
\input{Chapter-0/0-Introduction}
\mainmatter
\input{Chapter-1/1-Theory}
\input{Chapter-2/2-LArTPCs}
\input{Chapter-3/3-TheLArIATExperiment}
\input{Chapter-4/4-Methodology}
\input{Chapter-5/5-Preparatory}
\input{Chapter-6/6-PionCrossSection}
\input{Chapter-7/7-KaonCrossSection}
\input{Chapter-8/8-Conclusions}

% Only call appendix once, here.
\appendix
\input{AppendixB-EField/B-LArIATEField}
\input{AppendixB-TrackingStudies/B-Tracking}
\input{AppendixC-EnergyCalibration/C-EnergyCalibration}

% Add additional \chapter{}s as necessary.

% use \cite{} to cite a reference in your bibliography file.
% use \ref{} to reference a \label{} from an equation, figure, or table.

% for sets of equations use align:
%\begin{align}
%\end{align}

% for figures:
%\begin{figure}[ht]
%\centering
%\includegraphics[width=.45\textwidth]{name_of_figure.eps}
%\caption{A caption! \label{a_figure}}
%\end{figure}

% for tables:
%\begin{table}
%\begin{tabular}{c|c|c}
% 1 & 2 & 3 \\
%\hline
%\end{tabular}
%\caption{Another caption! \label{a_table}}
%\end{table}

% Any chapters such as End Notes go after this.
\backmatter

\bibliography{bib.bib}
%\bibliographystyle{plain}
% for your own sake, use a bibtex file, so all of the numbering of references will be done
% automatically.

\end{document}